---
title: "lab_02"
author: "derek willis"
date: "2023-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse) 
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? **A1. Atmospheric and Oceanic Science offered the most courses with climate in the title**

```{r}
umd_courses |>
  filter(str_detect(title, "Climate") & seats >= 1) |>
  group_by(department)|>
  summarize(classes = n()) |>
  arrange(desc(classes))
```

**Q2.** Of the departments listed in A2, which one is the most interesting to you, and why? **A2. Im interested why Germanic Studies is on the list, I can understand why some of the other majors would include the word climate in their course titles, but not Germanic Studies!**

**Q3** Which departments have the most courses in the Fall 2023 term? And which departments in that term have at least one syllabus posted? How are the top five departments different in each result?

**A3. Business and Management has the most courses in the Fall 2023 regardless of syllabus count. However,** Economics, Biological Sciences Program, Engineering, Professional Masters, and Public Policy have the most classes posted with a syllabus. The School of Music, English, TLPL, Info Studies, and Journalism all have more classes without a syllybus.

```{r}
umd_courses |>
  filter(term == 202308) |>
  group_by(department)|>
  summarize(courses = n()) |>
  arrange(desc(courses))

umd_courses |>
  filter(term == 202308 & syllabus_count > 1) |>
  group_by(department)|>
  summarize(courses = n()) |>
  arrange(desc(courses))
```

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr, saving the result as the same dataframe. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time. **A4. The spending increases as time increases, with an outlier of a dip from Sept to Nov**

```{r}
maryland_expenses |>
  mutate(expenditure_date = mdy(expenditure_date)) |>
  mutate(month = floor_date(expentditure_date, "month")) |>
  group_by(month) |>
  filter(str_detect(committee_name, "Moore Wes For Maryland")) |>
  summarise(total_amount = sum(amound)) |>
  arrange(month)
```

**Q5.** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record spent money in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for the others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe). **A5:**

```{r}
maryland_expenses |>
  mutate(
    spending_location = case_when(
      str_detect(address, "Maryland ") ~ "In-State",
        str_detect(address, "California ") ~ "Out of state",
      str_detect(address, " Washington ") ~ "Out of state",
      str_detect(address, " Louisiana ") ~ "Out of state",
      str_detect(address, " Florida ") ~ "Out of state",
      str_detect(address, " North Carolina ") ~ "Out of state",
      str_detect(address, " Massachusettes ") ~ "Out of state",
      str_detect(address, " West Virginia ") ~ "Out of state",
      str_detect(address, " Virginia ") ~ "Out of state",
      .default = NA
      
    )) |>
  group_by(spending_location) |>
  summarise(total_amount = sum(amount))
```

**Q6.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab. **A6.** Could you abbreviate class names?

```{r}
umd_courses <- umd_courses|>
  mutate(
    title = case_when(
        str_detect(class, "Journalism") ~ "Jour",
        str_detect(class, "Introductory Linguistics") ~ "Ling",
        .default = NA
    )
  )
```
