---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
homicide_data <- read.csv('data/bmore-homicide-data.csv') |>
view(homcide_data)
```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
homicide_data |>
  mutate(reported_date = ymd(reported_date)) |>
  mutate(year = year(reported_date)) |>
  filter(year == reported_date) |>
  
### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here] 
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
