---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(readr)
library(tidyverse)
library(janitor)
homicide_data <- read_csv("data/homicide_data.csv") 
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
homicide_data <- read.csv("data/homicide_data.csv") |>
view(homicide_data)

  homicide_data <- homicide_data |>
  mutate(fixed_date = ymd(reported_date), 
         year = year(fixed_date))
```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **"As Baltimore has seen a stunning surge of violence, with nearly a killing each day for the past three years in a city of 600,000, homicide arrests have plummeted."**
-   **We were able to take a look at the average per day, and the number of arrests. We found that what the article said was true, and we were also able to compare the difference between the three years prior, where killings were much less common.**

```{r}
#As Baltimore has seen a stunning surge of violence, with nearly a killing each day for the past three years in a city of 600,000, homicide arrests have plummeted.
killings_per_year <- homicide_data |>
  mutate(reported_date =ymd(reported_date)) |>
  mutate(year = year(reported_date)) |>
  filter(str_detect(city, 'Baltimore')) |>
  filter(year == '2017' | year == '2016' | year == '2015' | year == '2014' | year == '2013' | year == '2012') |>
  group_by(year) |>
  summarize(
    total_killings = n(),
    average_per_day = (total_killings/365)
)
```

```{r}
killings_per_year |>
  ggplot() +
  geom_bar(aes(x=year, weight=total_killings))+
  scale_x_continuous(breaks=c(2012, 2013, 2014, 2015, 2016, 2017)) +
  labs(
    title="Homicides Per Year in Balitmore",
    x = "Year",
    y = "Homicides Per Year",
    caption = "Source: The Washington Post"
  )+
  theme_clean()

```

### Sentence 2

-   **Of 50 of the nation's largest cities, Baltimore is one of 34 where police now make homicide arrests less often than in 2014, according to a Washington Post analysis.**
-   **We could only find 14 instead of 24, and aren't sure what the issue with that is. NY is difficult because they didn't provide data for all the years.**

```{r}
#Of 50 of the nation's largest cities, Baltimore is one of 34 where police now make homicide arrests less often than in 2014, according to a Washington Post analysis.
# Put code to reverse engineer sentence here
arrests_2014 <- homicide_data |>
  filter(year == '2014' & disposition == "Closed by arrest") |>
  group_by(city) |>
  summarize(arrests_2014 = n())

arrests_2017 <- homicide_data |>
  filter(year == '2017' & disposition == "Closed by arrest") |>
  group_by(city) |>
  summarize(arrests_2017 = n())

homicide_arrests <- arrests_2014|> 
  left_join(arrests_2017) |>
  mutate(pct_change = (arrests_2017 - arrests_2014)/arrests_2014) |>
  filter(pct_change < 0)


homicide_arrests
```

### Sentence 3

-   **Baltimore is also one of 30 cities that have seen an increase in homicides in recent years, with the greatest raw number increase in killings of any city other than Chicago, which has four times the population.**
-   We were able to group it to compare the difference in homicides between 2012 and 2017, which seem to be the years that the article was referring to.

```{r}
#Baltimore is also one of 30 cities that have seen an increase in homicides in recent years, with the greatest raw number increase in killings of any city other than Chicago, which has four times the population.
homicide_count_2012 <- homicide_data |>
  filter(year == "2012") |>
  group_by(city) |>
  summarize(homicide_count_2012 = n())

homicide_count_2017 <- homicide_data |>
  filter(year == "2017") |>
  group_by(city) |>
  summarize(homicide_count_2017 = n())

homicide_data_combined <- homicide_count_2017 |>
  left_join(homicide_count_2012) |>
  filter(city != "New York") |>
  mutate(raw_number_increase = homicide_count_2017 - homicide_count_2012) |>
  arrange(desc(raw_number_increase))

# Display results of code below this codeblock
homicide_data_combined
```

### Sentence 4

-   **Sentence text**: **For most of the decade before 2015, Baltimore's annual homicide arrest rate hovered at about 40 percent. Since 2015, the arrest rate hasn't topped 30 percent in any year.**

-   **Analysis summary**: We were able to confirm the finding. The arrest rate stayed below 30% after 2015, when it used to be much higher.

    ```{r}
    baltimore_arrests_by_year <- homicide_data |>
      filter(city == "Baltimore") |>
      filter(disposition == "Closed by arrest")|>
      group_by(year) |>
      summarize(arrests = n())

    baltimore_homicides_by_year <- homicide_data |>
      filter(city == "Baltimore") |>
      group_by(year) |>
      summarize(homicides = n())

    baltimore_arrest_rate <- baltimore_arrests_by_year|>
      left_join(baltimore_homicides_by_year, by = "year") |>
      mutate(arrest_rate = (arrests/homicides)*100)

    # Display results of code below this codeblock
    baltimore_arrest_rate
    ```

### Sentence 5

-   **Sentence text**: **Of the 1,002 homicides between 2015 and the beginning of this year, just 252 --- one out of every four --- resulted in an arrest.**

-   **Analysis summary**: We were able to confirm the finding. We found that there are roughly 1/4 arrests for every homicide from 2015-2018, with an exact percentage of 25.1497% of homicides resulting in arrests. This shows that there are less arrests occuring.

```{r}
recent_baltimore_homicides <- homicide_data |>
  filter(city == "Baltimore") |>
  filter(year == '2015' | year == '2016' | year == '2017' | year == '2018') 
  
total_arrest_count <- recent_baltimore_homicides|>
    filter(disposition == "Closed by arrest")
    

nrow(total_arrest_count)/nrow(recent_baltimore_homicides)*100
```

-30-
